---
title: "Lab_06"
output: html_document
date: "2022-10-17"
---
##Question 1
```{r}
require(palmerpenguins)

rm(list = ls())
sse_mean = function(x)
{
  get_sd = 
    function(x){
      return(sd(x, na.rm = TRUE))
    }
  
  get_sample_size =
    function(x){
      sample_size = sum(table(x))
      return(sample_size)
    }
  sse = get_sd(x) / sqrt(get_sample_size(x))
  return(sse)
}

sse_mean(penguins$body_mass_g)
sse_mean(mtcars$mpg)
dev.off()
```



```{r}
boxplot(
  flipper_length_mm ~ species, data = penguins,
  ylab = "Flipper length (mm)")

dat_pen = subset(penguins, species != "Chinstrap")
boxplot(
  flipper_length_mm ~ species, data = dat_pen,
  ylab = "Flipper length (mm)")

dat_pen = droplevels(subset(penguins, species != "Gentoo"))
{
  par(mfrow = c(1, 2))
  boxplot(
    flipper_length_mm ~ species, data = penguins,
    ylab = "Flipper length (mm)")
}

```
##Montecarlo distribution reshuffles all data values to simulate the null hypothesis
```{r}
# for reproducibility
set.seed(123)

flipper_shuffled = sample(
  penguins$flipper_length_mm, replace = TRUE)

{
  par(mfrow = c(1, 2))
  boxplot(
    flipper_length_mm ~ species, data = penguins,
    ylab = "Flipper length (mm)",
    main = "Original Data")
  boxplot(
    flipper_shuffled ~ penguins$species,
    ylab = "Flipper length (mm)",
    main = "MonteCarlo Resampled Data",
    xlab = "species")
}
```

##Bootstrapping to shuffle the data within the same columns
##Allows us to change the values within a column, comparing them to see how the data is associated
##Bootstrapping simulates an alternative hypothesis
## It shows the relationship between data but is not for the original data/experiment
```{r}
penguins2 = penguins[sample(1:nrow(penguins), replace = T), ]

{
  par(mfrow = c(1, 2))
  boxplot(
    flipper_length_mm ~ species, data = penguins,
    ylab = "Flipper length (mm)",
    main = "Original Data")
  boxplot(
    flipper_length_mm ~ species, data = penguins2,
    ylab = "Flipper length (mm)",
    main = "Bootstrap Data")
}
```
##Repeated MC tests
```{r}
par(mfrow = c(4, 4), mar = c(1, 1, 1, 1))
for (i in 1:16)
{
  
  flipper_shuffled = sample(
    penguins$flipper_length_mm, replace = TRUE)
  
  boxplot(
    flipper_shuffled ~ penguins$species,
    ann = F, axes = F)
  box()
  
}
```
```{r}
t.test(dat_pen$flipper_length_mm ~ dat_pen$species)

```

##  Welch Two Sample t-test

## data:  dat_pen$flipper_length_mm by dat_pen$species
## t = -5.7804, df = 119.68, p-value = 6.049e-08
## alternative hypothesis: true difference in means between group Adelie and group Chinstrap is not equal to 0
## 95 percent confidence interval:
##  -7.880530 -3.859244
## sample estimates:
##    mean in group Adelie mean in group Chinstrap 
##                189.9536                195.8235

```{r}
agg_means = aggregate(
  flipper_length_mm ~ species, 
  data = dat_pen, 
  FUN = "mean", 
  na.rm = TRUE)
diff_observed = diff(agg_means[, 2])

agg_means
diff_observed
```

```{r}
table(dat_pen$species)

```
```{r}
n_1 = 68
n_2 = 152

dat_1 = sample(dat_pen$flipper_length_mm, n_1, replace = TRUE)
dat_2 = sample(dat_pen$flipper_length_mm, n_2, replace = TRUE)

diff_simulated = 
  mean(dat_1, na.rm = TRUE) - mean(dat_2, na.rm = TRUE)
diff_simulated

print(c(observed = diff_observed, simulated = diff_simulated))
```

```{r}
x = dat_pen$flipper_length_mm
n_1 = 68
n_2 = 152

dat_1 = sample(x, n_1, replace = TRUE)
dat_2 = sample(x, n_2, replace = TRUE)

diff_simulated = 
  mean(dat_1, na.rm = TRUE) - mean(dat_2, na.rm = TRUE)
```

##creating the resampling function
```{r}
two_group_resample_diff = function(x, n_1, n_2){
  dat_1 = sample(na.omit(x), n_1, replace = TRUE)
  dat_2 = sample(na.omit(x), n_2, replace = TRUE)

  difference_in_means = mean(dat_1, na.rm = TRUE) - mean(dat_2, na.rm = TRUE)
  
  return(difference_in_means)
}

set.seed(54321)
two_group_resample_diff(dat_pen$flipper_length_mm, 68, 152)
```
##Resampling flipper lengths
##Question 4
```{r}
n = 2000
mean_differences = c()
for (i in 1:n)
{
  mean_differences = c(
    mean_differences,
    two_group_resample_diff(dat_pen$flipper_length_mm, 68, 152)
  )
}
hist(mean_differences)
```
```{r}
sum(abs(mean_differences) > diff_observed)
```

##Questions 7-11
```{r}
boxplot(
  body_mass_g ~ species, data = penguins,
  ylab = "Body Mass (g)", main = "Question 7")
```
##Question 8
```{r}
agg_means = aggregate(
  body_mass_g ~ species, 
  data = dat_pen, 
  FUN = "mean", 
  na.rm = TRUE)
diff_crit = diff(agg_means[, 2])

agg_means
diff_crit
```
##Question 9
```{r}
t.test(dat_pen$body_mass_g ~ dat_pen$species)
```
##Question 10/11
```{r}
table(dat_pen$species)
n = 1000
mean_differences = c()
for (i in 1:n)
{
  mean_differences = c(
    mean_differences,
    two_group_resample_diff(dat_pen$body_mass_g, 152, 124)
  )
}
hist(mean_differences, main = "Question 11")
sum(abs(mean_differences) > diff_crit)
```

